---
title: "Visualizations of the American Food System"
author: "Julia Whiting"
output: 
  html_document:
    code_folding: hide
---

```{r setup, echo=FALSE,warning=FALSE, message=FALSE, error=FALSE}
library(GISTools)
library(treemapify)
library(maps)
library(waffle)
library(tidyverse)
library(extrafont)
library(extrafontdb)
library(rgeos)
library(haven)
library(readxl)
library(rmarkdown)
library(ggplot2)
library(forcats)
library(stringr)
library(stringi)
library(hexbin)
library(scales)
library(cartogram)
library(geojsonio)
library(RColorBrewer)
library(broom)
library(sf)
library(data.tree)
library(treemap)
library(ggmosaic)
library(showtext)
```

```{r echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
font_add_google("Dosis", family = "dosis", bold.wt = 700)
font_add_google("Open Sans", family = "open")
showtext.auto()
main_theme <- theme(plot.title = element_text(color = "black",
                                              family = "dosis",
                                              hjust = 0.5,
                                              size = 14),
                    plot.subtitle = element_text(color = "#000000",
                                                 family = "dosis",
                                                 hjust = 0.5,
                                                 face = "italic",
                                                 size = 12),
                    plot.caption = element_text(size = 7,
                                                family = "dosis",
                                                face = "italic"),
                    axis.title.y = element_text(size = 11,
                                                family = "dosis"),
                    axis.title.x = element_text(size = 11,
                                                family = "dosis"),
                    axis.text.x = element_text(family = "dosis"),
                    axis.text.y = element_text(family = "dosis"),
                    axis.line = element_line(color = "#000000"),
                    panel.background = element_blank(),
                    panel.grid.major = element_blank(),
                    legend.background = element_rect(fill = "transparent"),
                    legend.direction = "horizontal",
                    legend.text = element_text(size = 7,
                                               family = "dosis"),
                    legend.position = "top",
                    legend.title = element_text(size = 9,
                                                family = "dosis"))

map_theme <-theme(plot.title = element_text(color = "black",
                                              family = "Tahoma Bold",
                                              hjust = 0.5,
                                              size = 14),
                    plot.subtitle = element_text(color = "#000000",
                                                 family = "Tahoma",
                                                 hjust = 0.5,
                                                 face = "italic"),
                    plot.caption = element_text(size = 9,
                                                family = "Tahoma",
                                                face = "italic"),
                    legend.position = c(0.5, 0.9),
                    panel.background = element_blank(),
                    axis.title.x  = element_blank(),
                    axis.title.y = element_blank(),
                    axis.text.x = element_blank(),
                    axis.text.y = element_blank(),
                    axis.ticks = element_blank(),
                    axis.line = element_blank(),
                    legend.background = element_rect(fill = "transparent"),
                    legend.direction = "horizontal",
                    legend.text = element_text(size = 7),
                    legend.title = element_text(size = 9))
  

categorical_cols <- c( "#006D2F","#FF8900", "#907EAE", "#054361", "#DAAC35", "#992000", "#0E1969")
alt_cols <- c('#006d2f','#49742c','#737b27','#968122','#b9851b','#db8812','#ff8900')
diverging_grad <- c('#006d2f','#5c782a','#8c8224','#b58c1d','#d99914','#f3a908','#ffc100')
alt_diverg <- c('#006d2f','#6c7a28','#ae841e','#ed890a','#ffa500','#ffcb1b')
alt2_diverg <- c('#ffcb1b','#f3a908','#ed890a','#6c7a28','#ae841e', '#006d2f')
sml_diverg <- c('#006d2f','#687929','#a5831f','#e3880f','#ff9d00','#ffc100')
smol_diverg <- c('#ffc100','#ff9d00', '#a5831f','#687929','#006d2f')
```
<style type="text/css">

h1.title {
  font-size: 28px;
  font-family: 'Open Sans',sans-serif
  color: Black;
  text-align: center;
}
h4.author {   Header 4 - and the author and data headers use this too   
    font-size: 14px;
  font-family: 'Open Sans', sans-serif;
  color: Black;
  text-align: center;
}
body{
  font-family: 'Dosis', sans-serif;
  font-size: 10pt;
}
</style>
```{r warning=FALSE, message=FALSE, error=FALSE,echo=FALSE, fig.showtext=TRUE, fig.align="center"}
farmVals <- read_xlsx("farmvals.xlsx")
font_add_google("Dosis", family = "dosis")
font_add_google("Open Sans", family = "open")
showtext_auto(enable = TRUE)
all_values <- farmVals %>%
  mutate(farm_value = fct_collapse(farm_value,
                                   "< $4999" = c("<$2500", "$2500-4999")),
         farm_value = factor(farm_value, levels = c("< $4999",
                                                    "$5000-9999",
                                                    "$10,000-24,999",
                                                    "$25,000-49,999",
                                                    "$50,000-99,999",
                                                    "$100,000+"))) %>%
  group_by(farm_value) %>%
  summarise(tot_farms = sum(total_farms),
            proportion = tot_farms/sum(tot_farms),
            ymax = cumsum(proportion),
            ymin = c(0, head(ymax, n=-1))) 
showtext.begin()
all_values %>%
  ggplot(aes(x = farm_value, fill = farm_value, y = tot_farms)) +
    geom_col() +
    main_theme +
    labs(title = "Farming in America is a business of extremes",
         subtitle = "Most farms' values in sales is either under $4999 or above $100,000",
         caption = "Source: United States Agricultural Data, https://doi.org/10.3886/ICPSR35206.v4",
         x = "Value of Farms",
         y = "Number of Farms",
         fill = "Value of Farms\nby Sales") +
    scale_fill_manual(values = sml_diverg) +
    theme(legend.position = "none") +
    scale_y_continuous(expand=c(0,0),
                       labels = comma) 

smal_diverg <- c('#006d2f','#ffc100', '#a5831f','#687929','#e3880f','#ff9d00')
```



Taking annual sales to stand in for farm value, an interesting pattern emerges.
Most farms in the United States earn less than $5000. The next most are the top earners,
taking in over $100,000 in sales. It would appear that farming as a profession 
with a reasonable income no longer the norm. 






```{r warning=FALSE, message=FALSE, error=FALSE, echo=FALSE, fig.showtext=TRUE,fig.align="center"}
access <- read_xlsx("foodAtlas.xlsx", sheet = 5)
assistance <- read_xlsx("foodAtlas.xlsx", sheet = 6)
security <- read_xlsx("foodAtlas.xlsx", sheet = 8)
health <- read_xlsx("foodAtlas.xlsx", sheet = 7)
farmsTime <- read_xlsx("farmnumbers.xlsx")
agCensus <- read_csv("agcensus.csv")
access <- access %>%
  mutate(FIPS = as.integer(FIPS))
assistance <- assistance %>%
  mutate(FIPS = as.integer(FIPS))
security <- security %>%
  mutate(FIPS = as.integer(FIPS))
health <- mutate(health, FIPS = as.integer(FIPS))

accAss <- right_join(access, assistance, by = "FIPS")
accAssSec <- right_join(accAss, security, by = "FIPS")
ctyFarms <- inner_join(accAssSec, agCensus, by = "FIPS")
ctyFarms <- inner_join(ctyFarms, health, by = "FIPS") %>%
              filter(COUNICP != 0)
farm_payments <- ctyFarms %>%
  group_by(State.x) %>%
  summarise(payments = sum(DATA1_33, na.rm = TRUE))
#code for hex map from: https://www.r-graph-gallery.com/328-hexbin-map-of-the-usa/
# Hexbin available in the geojson format here: https://team.carto.com/u/andrew/tables/andrew.us_states_hexgrid/public/map. Download it and then:
spdf <- geojson_read("us_states_hexgrid.geojson.json",  what = "sp")

farm_payments <- read_xlsx("farmpayments.xlsx")%>%
  mutate(Payments = Payments/1000)


spdf@data = spdf@data %>% mutate(google_name = gsub(" \\(United States\\)", "", google_name))
spdf_fortified <- tidy(spdf, region = "google_name")

centers <- cbind.data.frame(data.frame(gCentroid(spdf, byid=TRUE), id=spdf@data$iso3166_2))
centers <- centers %>% filter(id != "DC")

spdf_fortified = spdf_fortified %>%
  left_join(. , farm_payments, by=c("id"="State")) 

showtext.begin()

spdf_fortified %>%
  filter(id != "District of Columbia") %>%
  ggplot() +
    geom_polygon(aes(fill = Payments, x = long, y = lat, group = group),
                 size=0.05, alpha=.8, color="black") +
    main_theme +
    coord_map() +
    geom_text(data=centers, aes(x=x, y=y, label=id), color="black", size=3, family = "dosis") +
    scale_fill_gradientn(colours=brewer.pal(7, "Greens"),
                         name="Government Payments to Farms (mil)",
                         breaks = seq(0, 800, 100)) +
  theme(legend.position = c(0.5, 0.9),
        legend.text = element_text(angle = 45,
                                   family = "dosis"),
    plot.background = element_blank(),
    axis.title.x  = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank()) +
  labs(title = "Iowa, Illinois, and Texas recieved one quarter of all\ngoverment payments to farms in 2012.",
       caption = "Source: United States Agricultural Data, https://doi.org/10.3886/ICPSR35206.v4",
       subtitle = "Midwestern and plains states received the most subsidy payments, and New England\n received the fewest.")
```



Many farms receive payments from the government through subsidies. Most 
subsidies fund commodity crops, such as corn, wheat, and cotton. Subsidies are 
concentrated in states that produce these commodities, mostly in the Midwest 
and Plains states. Because of the relationship between crops and subsidies, 
the state that receives the most government payments is not the state with the 
most farms.





```{r warning=FALSE, message=FALSE, error=FALSE, echo=FALSE, fig.showtext=TRUE, fig.align="center"}
frm_size <- read_xlsx("farmsize.xlsx")

showtext.begin()

frm_size %>%
  mutate(State.x = as.factor(State.x),
         State.x = fct_reorder(State.x, -all_farms),
         farm_size = as.factor(farm_size),
         farm_size = fct_collapse(farm_size,
                                  "10-179 acres" = c("10-49 Acres", "50-179 Acres")))%>%
  ggplot(aes(x = State.x, y = total_farms)) +
    geom_col(aes(fill = farm_size)) + 
    labs(x = "State",
         y = "Number of Farms",
         title = "American agriculture is an industry of small farms",
         subtitle = "Most farms are between 10-180 acres nationwide.",
         caption = "Data Source: United States Agricultural Data, https://doi.org/10.3886/ICPSR35206.v4",
         fill = "Farm Size") +
    main_theme +
    theme(axis.text.x = element_text(angle = 90),
          legend.position = c(0.55, 0.8))+ 
  scale_y_continuous(expand = c(0, 0),
                     labels = comma) +
  scale_fill_manual(values = smol_diverg, aesthetics = "fill")
```



Most farms in the US are relatively small. Many food experts will tell you
that farm sizes in the US are growing. Yet most farms are still under .25 square
miles (180 acres). Texas has approximately double the number of farms than the 
next state.





```{r warning=FALSE, message=FALSE, error=FALSE, echo=FALSE, fig.showtext=TRUE, fig.align="center"}
byState <- ctyFarms %>%
  group_by(State.x) %>%
  summarise(low_acc_pop = sum(LACCESS_POP10)/1000000,
            farms = sum(DATA1_1),
            snap_redemp = sum(REDEMP_SNAPS12,
                              na.rm = TRUE)/1000000,
            payments = sum(DATA1_33, na.rm = TRUE),
            cropland = sum(DATA1_15, na.rm = TRUE),
            farmland = sum(DATA1_2, na.rm = TRUE),
            avg_land = mean(DATA1_3, na.rm = TRUE),
            obeseAd = mean(PCT_OBESE_ADULTS13),
            insec = mean(VLFOODSEC_10_12 + FOODINSEC_10_12)) %>%
  filter(State.x != "0")
local <- read_xlsx("foodAtlas.xlsx", sheet = 9)
local <- local %>%
  mutate(FIPS = as.integer(FIPS))
accAss <- right_join(access, assistance, by = "FIPS")
accAssSec <- right_join(accAss, security, by = "FIPS")
ctyFarms <- inner_join(accAssSec, agCensus, by = "FIPS")
ctyFarms <- inner_join(ctyFarms, local, by = "FIPS") %>%
              filter('COUNICP' != 0)
showtext_begin()

byState %>%
  ggplot(aes(x = farms, y = low_acc_pop)) +
    geom_point(aes(color = insec), 
               alpha = 0.9, 
               size = 7) +
    geom_text(aes(label = State.x),size = 2, color = "white", family = "dosis") +
    scale_size_continuous(range = c(3, 8), labels = comma) +
    scale_y_continuous(breaks = pretty(byState$low_acc_pop)) +
    scale_x_log10(breaks = c(1000, 2500, 5000, 10000, 15000, 20000,25000,
                             50000,75000, 100000, 150000, 200000, 250000))+
    scale_color_continuous(low = "#e5f5e0", 
                           high = "#006d2c", 
                           guide = "colorbar",
                           breaks = seq(0,28,4)) +
  labs(x = "Number of Farms (log)",
       y = "Residents with Low Food Access (Millions)", 
       color = "Percent Food\nInsecure",
       caption = "Data Source: Food Environment Atlas, United States Agricultural Data",
         title = "Larger states have high numbers of farms and residents with\nlow food access",
         subtitle = "Food security is not related to state food production, but low access is positively related to number of \nfarms.") +
  main_theme +
  theme(legend.position = c(0.2,0.6), 
        legend.background = element_rect(fill = "transparent"),
        legend.title = element_text(size = 9),
        axis.text.x = element_text(angle = 90)) +

  stat_smooth(method="lm", se=FALSE, 
                formula=(y ~ exp(x)),colour="#ffa500", linetype = "dashed") 
```



There is a positive relationship between number of farms in a state and number of 
residents with low food access. Low access to food is defined as living more than one mile
away from a groccery store in urban areas, and more than 10 miles away from a 
groccery store in rural areas. Food insecurity is defined as being unable to 
provide adequate food for one or more household members at some time during 
the year. While food access is related to number of farms, food insecurity is not
related to access to food. There may be food produced in a state, but the ability of 
residents to buy that food is another matter entirely.  





```{r warning=FALSE, message=FALSE, error=FALSE, echo = FALSE, fig.showtext=TRUE, fig.align="center"}
crops <- agCensus %>%
  select(COUNICP, NAME,DATA1_16, DATA1_139, DATA1_133, DATA1_64, DATA1_68, DATA1_70, DATA1_94,
         DATA1_103, DATA1_82, DATA1_85, DATA1_88, DATA1_97, DATA1_109, DATA1_112, DATA1_115,
         DATA1_121, DATA1_124, DATA1_127, DATA1_130, DATA1_91) %>%
  filter(COUNICP == 0) %>%
  head(n = 50) 

crop_name <- crops %>%
  gather(DATA1_139, DATA1_133, DATA1_64, DATA1_68, DATA1_70, DATA1_94,
         DATA1_103, DATA1_82, DATA1_85, DATA1_88, DATA1_97, DATA1_109, DATA1_112, DATA1_115,
         DATA1_121, DATA1_124, DATA1_127, DATA1_130, DATA1_91,
         key = "crop", value = "acres") %>%
  mutate(crop = as.factor(crop)) %>%
  mutate(crop = fct_recode(crop,
                            "Orchards"="DATA1_139",
                            "Vegetables"="DATA1_133" ,
                            "Corn" = "DATA1_64",
                            "Silage corn"= "DATA1_68",
                            "Wheat" = "DATA1_70",
                            "Soybeans" = "DATA1_94" ,
                            "Cotton" = "DATA1_103"  ,
                            "Oats" = "DATA1_82" ,
                            "Barley" = "DATA1_85"  ,
                            "Sorghum" = "DATA1_88" ,
                            "Sorghum silage" = "DATA1_91",
                            "Dry Beans" = "DATA1_97"  ,
                            "Tobacco" = "DATA1_109"  ,
                            "Hay" = "DATA1_112"  ,
                            "Rice" = "DATA1_115" ,
                            "Sugar Beets" = "DATA1_121" ,
                            "Sugar Cane" = "DATA1_124"  ,
                            "Pineapples" = "DATA1_127" ,
                            "Peanuts" = "DATA1_130")) %>%
  na.omit() 

percents <-
crop_name %>%
  mutate(crop = fct_collapse(crop,
                             "Hay and Silage" = c("Silage corn",
                                               "Hay",
                                               "Sorghum silage"),
                             "Other" = c("Peanuts",
                                         "Sugar Cane",
                                        "Sugar Beets",
                                        "Dry Beans",
                                        "Tobacco",
                                        "Cotton",
                                        "Oats",
                                         "Barley",
                                         "Sorghum",
                                         "Rice"),
                             "Fruits/Vegetables" = c("Vegetables",
                                                     "Pineapples",
                                                     "Orchards"))) %>%
  select(acres, crop)%>%
  mutate(crop = factor(crop,levels = c("Hay and Silage",
                       "Corn", 
                       "Soybeans",
                       "Wheat",
                       "Other",
                       "Fruits/Vegetables"))) %>%
  group_by(crop) %>%
  summarise(tot_acres = sum(acres),
            proportion = tot_acres/389690400)  
showtext_begin()
parts <- c('Hay and Silage' = (100 - 22 - 20 - 13 - 7 - 2), 'Corn' = 22, 'Soybeans' = 20,
           'Wheat' = 13, 'Other' = 7, 'Fruits Vegetables' = 2)
waffle(parts, rows = 10, size = 1, colors = alt2_diverg) +
  labs(title = "The majority of cropland in the US is used to feed animals,\nnot people.",
       subtitle = "1 Square = 1% of US Cropland",
       caption = "Source: United States Agricultural Data, https://doi.org/10.3886/ICPSR35206.v4") +
  main_theme +
  theme(axis.line = element_blank(),
        legend.position = "right",
        legend.direction = "vertical") +
  annotate("text", x = 2.5, y = 5, label = "Hay and silage\ngrown for\n animal feed", family = "dosis", fontface = 2) +
  geom_rect(xmin = 6.5, xmax = 8.5, ymin = 0.5, ymax=7.5,  color = "#000000", fill=alpha("grey",0)) +
  annotate("text", x = 7.5, y = 5.7, label =  "Soy\ngrown\nfor\nanimal\nfeed", family = "dosis", fontface = 2) +
  geom_rect(xmin=4.5, xmax=6.5,ymin=0.5,ymax=5.2,color="#000000", fill = alpha("grey",0)) +
  annotate("text",x=5.5,y=3.7,label="Corn\ngrown\nfor\nanimal\nfeed", family="dosis", fontface=2)
```



The majority of crop land in the United States is not used to feed people. Hay and silage, both grown for animal feed, account for 43% of all US cropland. Additionally, 48% of corn and 70% of soy is grown to produce animal feed. In total, approximately 65% of cultivated land in the country produces food not fit for human consumption. In contrast, fruits and vegetables take up just 2% of cropland in the country.  





```{r warning=FALSE, message=FALSE, error=FALSE, echo=FALSE, fig.showtext=TRUE, fig.align="center"}
mkts <- ctyFarms %>%
  group_by(State.x) %>%
  summarise(Markets = sum(FMRKT13, na.rm = TRUE),
            "SNAP Markets" = sum(FMRKT_SNAP13, na.rm = TRUE)) %>%
  gather(Markets, "SNAP Markets", 
         key = "type", value = "number") %>%
  filter(State.x != 0)
showtext_begin()
ggplot(data = mkts, aes(area = number, fill = type, subgroup = State.x),
       alpha = 0.2) +
  geom_treemap() +
  scale_fill_manual(values = c('#5c782a',"#f3a908")) +
  geom_treemap_subgroup_text(color = "White", size = "10", place = "center", family = "dosis") +
  geom_treemap_subgroup_border(color = "White") +
  labs(title = "Coastal states have more farmers' markets\n that accept SNAP benefits",
       subtitle = "Illinois, Oregon, Vermont, and Michigan have high proportions of farmers markets\n that accept SNAP",
       caption = "Source: Food Environent Atlas, https://data.world/adamhelsinger/food-environment-atlas",
       fill = "Type of Market") +
  main_theme +
  theme(legend.position = "right",
        legend.direction = "vertical")
  
```



Farmers markets remain concentrated on the coasts. States in New England stand out
as having especially high proportions of markets that accept SNAP benefits. California
and New York have the most farmers markets in the nation, which makes sense given their 
populations. Michigan and Illinois have large numbers of farmers' markets and markets 
that accept SNAP benefits given their populations. 
\n
\n
\n


```{r warning=FALSE, message=FALSE, error=FALSE, echo=FALSE, fig.showtext=TRUE, fig.align="center", fig.height=7,fig.width=8.5}
timefarms <- read_xlsx("timeseris.xlsx") 


timefarms %>%
ggplot(aes(x = year, y = Rank)) +
  geom_line(aes(group = Name), color = "light grey", alpha = 0.7) +
  geom_line(data = timefarms %>% filter(Name == "MS"), color = "#ffc100", alpha = .7) +
  geom_line(data = timefarms %>% filter(Name == "NC"), color = "#ffc100", alpha = .7) +
  geom_line(data = timefarms %>% filter(Name == "AK"), color = "#ffc100", alpha = .7) +
  geom_line(data = timefarms %>% filter(Name == "AL"), color = "#ffc100", alpha = .7) +
  geom_line(data = timefarms %>% filter(Name == "TX"), color = "#006d2f", alpha = .5) +
  geom_line(data = timefarms %>% filter(Name == "OK"), color = "#006d2f", alpha = .5) +
  geom_line(data = timefarms %>% filter(Name == "CA"), color = "#006d2f", alpha = .5) +
  geom_line(data = timefarms %>% filter(Name == "FL"), color = "#006d2f", alpha = .5) +
  theme(legend.position = "none") +
  main_theme +
  scale_x_continuous(breaks = c(1950, 1959, 1969, 1974, 1979, 1987, 1992, 1997, 2002, 2007, 2012),
                     expand = expand_scale(add = 1)) +
  scale_y_reverse(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15,
                                16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29,
                                30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 
                                42, 43, 44, 45, 46, 47, 48, 49, 50),
                     expand = c(0, 1), 
                     sec.axis = dup_axis()) +
  theme(legend.position = "none") +
  geom_text(aes(label = Name), size = 3, family = "dosis", color = "light grey", alpha = 0.7) +
  geom_text(data=timefarms%>%filter(Name=="TX"), aes(label = Name),
            color = "#006d2f", family = "dosis", size = 4, fontface=2, alpha = 0.7) +
  geom_text(data=timefarms%>%filter(Name=="CA"), aes(label = Name),
            color = "#006d2f", family = "dosis", size = 4, fontface=2, alpha = 0.7) +
  geom_text(data=timefarms%>%filter(Name=="FL"), aes(label = Name),
            color = "#006d2f", family = "dosis", size = 4, fontface=2, alpha = 0.7) +
  geom_text(data=timefarms%>%filter(Name=="OK"), aes(label = Name),
            color = "#006d2f", family = "dosis", size = 4, fontface=2, alpha = 0.7) +
  geom_text(data=timefarms%>%filter(Name=="NC"), aes(label = Name),
            color = "#ffc100", family = "dosis", size = 4, fontface=2) +
  geom_text(data=timefarms%>%filter(Name=="MS"), aes(label = Name),
            color = "#ffc100", family = "dosis", size = 4, fontface=2) +
  geom_text(data=timefarms%>%filter(Name=="AL"), aes(label = Name),
            color = "#ffc100", family = "dosis", size = 4, fontface=2) +
  geom_text(data=timefarms%>%filter(Name=="AK"), aes(label = Name),
            color = "#ffc100", family = "dosis", size = 4, fontface=2) +
  labs(x = "Year",
       y = "Rank of Number of Farms",
       subtitle = "North Carolina, Mississippi, and Alabama had the largest losses in farms,\n Oklahoma, California, and Florida had the largest gains",
       title = "Texas has had the most farms in the country, and Alaska the fewest, for 62 years",
       caption = "Source: Source: United States Agricultural Data, https://doi.org/10.3886/ICPSR35206.v4")

```



When looking at the ranking of states by number of farms from 1950-2012, several trends are apparent. Texas and Alaska have had the most and least farms, respectively, for the entire 62 year span. Given the size of Texas and geographical location of Alaska, their positions are not surprising. The states that saw the largest falls in rankings are southern: South Carolina, North Carolina, Mississippi, and Alabama. The states with the largest gains are more dispersed: Oklahoma, Florida, Colorado, and California. 
\n
\n
\n


```{r warning=FALSE, message=FALSE, error=FALSE, echo=FALSE, fig.showtext=TRUE, fig.align="center"}
number <- read_xlsx("byyear.xlsx")
showtext_begin()
number %>%
  ggplot(aes(x = year, y = farms)) +
  geom_line(aes(group=NAME), color = "light grey") +
  main_theme +
  scale_y_continuous(expand = c(0,0), 
                     labels = comma) +
  scale_x_continuous(breaks = c(1950, 1959, 1969, 1974, 1979, 1987, 1992, 1997, 2002, 2007, 2012),
                    expand = c(0,.4)) +
  geom_line(data = number %>%filter(NAME == "Texas"), color = "#f3a908") +
  geom_line(data = number %>%filter(NAME == "California"), color = "#5c782a") +
  annotate("text", x = 1982, y = 200000, label = "Texas", family = "dosis")+ 
  annotate("text", x = 1974, y = 85000, label = "California", family = "dosis") +
  
  annotate("text", x = 1998, y = 146000, label = "California, which changed from 21st\nmost agricultural to 5th most in this period,\n has more than 3x fewer farms as Texas", family = "dosis") +
  annotate("text", x=1980, y = 305000, label = "The gap in number of farms between Texas\n and other states has grown from 1950-2012",
           family = "dosis", size = 5) +
  labs(x = "Year",
       y = "Farms",
       caption = "Source: United States Agricultural Data, https://doi.org/10.3886/ICPSR35206.v4")
```



Almost every state lost farms between 1950 and 2012. Texas stands out for several reasons. It has more farms than any other state, which is not necessarily surprising given its area. What is unusual, however,
is that Texas has gained a large number of farms during a time when other states experienced losses, or extremely moderate gains. California, for example, went from being the 21st state with the most farms to 5th most from 1950-2012. This sounds impressive, but looking at the actual numbe of farms in the state, it becomes apparent that California's shift in rankings is mostly due to other states losing farms. In 1950, the state had 137,168 farms, while in 2012 it had 77,857. 





Data Sources:  
 Haines, Michael, Fishback, Price, and Rhode, Paul. United States Agriculture Data, 1840 - 2012. Ann Arbor, MI: Inter-university Consortium for Political and Social Research [distributor], 2018-08-20. https://doi.org/10.3886/ICPSR35206.v4 

 Helsinger, Adam. United States Department of Agriculture Food Environment Atlas. 2017. https://data.world/adamhelsinger/food-environment-atlas 