---
title: "Static Style"
author: "Julia Whiting"
date: "2/3/2019"
output: html_document
code_folding: hide
---

```{r setup, echo=FALSE,warning=FALSE, message=FALSE, error=FALSE}
library(tidyverse)
library(extrafont)
library(extrafontdb)
library(rgeos)
library(haven)
library(readxl)
library(rmarkdown)
library(ggplot2)
library(forcats)
library(stringr)
library(stringi)
library(hexbin)
library(scales)
library(cartogram)
library(geojsonio)
library(RColorBrewer)
library(broom)
```

## Theme Work

```{r warning=FALSE, message=FALSE, error=FALSE, echo = FALSE}
font_import()
fonts()
fonttable()

main_theme <- theme(plot.title = element_text(color = "black",
                                              family = "Tahoma Bold",
                                              hjust = 0.5,
                                              size = 12),
                    plot.subtitle = element_text(color = "#000000",
                                                 family = "Tahoma",
                                                 hjust = 0.5),
                    plot.caption = element_text(size = 7,
                                                family = "Tahoma"),
                    axis.title.y = element_text(size = 11,
                                                family = "Tahoma"),
                    axis.title.x = element_text(size = 11,
                                                family = "Tahoma"),
                    axis.line = element_line(color = "#000000"),
                    panel.background = element_blank(),
                    panel.grid.major = element_blank(),
                    legend.background = element_rect(fill = "transparent"),
                    legend.direction = "horizontal",
                    legend.text = element_text(size = 7),
                    legend.position = "top",
                    legend.title = element_text(size = 9))


categorical_cols <- c( "#FF8900", "#907EAE", "#054361", "#DAAC35", "#992000","#006D2F", "#0E1969")
```

## Farm Values In Sales

```{r warning=FALSE, message=FALSE, error=FALSE}
farmVals <- read_xlsx("farmvals.xlsx")

all_values <- farmVals %>%
  mutate(farm_value = fct_collapse(farm_value,
                                   "< $4999" = c("<$2500", "$2500-4999")),
         farm_value = factor(farm_value, levels = c("< $4999",
                                                    "$100,000+",
                                                    "$10,000-24,999",
                                                    "$5000-9999",
                                                    "$25,000-49,999",
                                                    "$50,000-99,999"))) %>%
  group_by(farm_value) %>%
  summarise(tot_farms = sum(total_farms),
            proportion = tot_farms/sum(tot_farms),
            ymax = cumsum(proportion),
            ymin = c(0, head(ymax, n=-1))) 

all_values %>%
  ggplot(aes(x = farm_value, fill = farm_value, y = tot_farms)) +
    geom_col() +
    main_theme +
    labs(title = "Farming in America is a business of extremes",
         subtitle = "Most farms' value in sales is either under $4999 or above $100,000.",
         caption = "Source = 2012 Agricultural Census",
         x = "Value of Farms",
         y = "Number of Farms",
         fill = "Value of Farms\nby Sales") +
    scale_fill_manual(values = categorical_cols, aesthetics = "fill") +
    theme(legend.position = "none")
```


## Farm Payments

Many farms receive payments from the government through subsidies. Most 
subsidies fund commodity crops, such as corn, wheat, and cotton. Subsidies are 
concentrated in states that produce these commodities, mostly in the Midwest 
and Plains states. Because of the relationship between crops and subsidies, 
the state that receives the most government payments is not the state with the 
most farms. 

```{r warning=FALSE, message=FALSE, error=FALSE}
access <- read_xlsx("foodAtlas.xlsx", sheet = 5)
assistance <- read_xlsx("foodAtlas.xlsx", sheet = 6)
security <- read_xlsx("foodAtlas.xlsx", sheet = 8)
health <- read_xlsx("foodAtlas.xlsx", sheet = 7)
farmsTime <- read_xlsx("farmnumbers.xlsx")
agCensus <- read_csv("agcensus.csv")
access <- access %>%
  mutate(FIPS = as.integer(FIPS))
assistance <- assistance %>%
  mutate(FIPS = as.integer(FIPS))
security <- security %>%
  mutate(FIPS = as.integer(FIPS))
health <- mutate(health, FIPS = as.integer(FIPS))

accAss <- right_join(access, assistance, by = "FIPS")
accAssSec <- right_join(accAss, security, by = "FIPS")
ctyFarms <- inner_join(accAssSec, agCensus, by = "FIPS")
ctyFarms <- inner_join(ctyFarms, health, by = "FIPS") %>%
              filter(COUNICP != 0)
farm_payments <- ctyFarms %>%
  group_by(State.x) %>%
  summarise(payments = sum(DATA1_33, na.rm = TRUE))
#code for hex map from: https://www.r-graph-gallery.com/328-hexbin-map-of-the-usa/
# Hexbin available in the geojson format here: https://team.carto.com/u/andrew/tables/andrew.us_states_hexgrid/public/map. Download it and then:
spdf <- geojson_read("us_states_hexgrid.geojson.json",  what = "sp")

farm_payments <- read_xlsx("farmpayments.xlsx")%>%
  mutate(Payments = Payments/1000)


spdf@data = spdf@data %>% mutate(google_name = gsub(" \\(United States\\)", "", google_name))
spdf_fortified <- tidy(spdf, region = "google_name")

centers <- cbind.data.frame(data.frame(gCentroid(spdf, byid=TRUE), id=spdf@data$iso3166_2))
centers <- centers %>% filter(id != "DC")

spdf_fortified = spdf_fortified %>%
  left_join(. , farm_payments, by=c("id"="State")) 

spdf_fortified %>%
  filter(id != "District of Columbia") %>%
  ggplot() +
    geom_polygon(aes(fill = Payments, x = long, y = lat, group = group),
                 size=0.05, alpha=0.6, color="black") +
    main_theme +
    coord_map() +
    geom_text(data=centers, aes(x=x, y=y, label=id), color="black", size=3) +
    scale_fill_gradientn(colours=brewer.pal(10, "Greens"),
                         name="Government Payments to Farms (mil)",
                         labels = comma) +
  theme(legend.position = c(0.5, 0.9),
    plot.background = element_blank(),
    axis.title.x  = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()) +
  labs(title = "Iowa, Illinois, and Texas recieved one quarter of all\ngoverment payments to farms in 2012.",
       caption = "Source: 2012 Agricultural Census",
       subtitle = "Midwestern and plains states received the most subsidy payments, and New England\n received the fewest.")
```

## Farm Size

Most farms in the US are relatively small. Many food experts will tell you
that farm sizes in the US are growing. Yet most farms are still under .25 square
miles (180 acres). Texas has approximately double the number of farms than the 
next state.

```{r warning=FALSE, message=FALSE, error=FALSE}
frm_size <- read_xlsx("farmsize.xlsx")

frm_size %>%
  mutate(State.x = as.factor(State.x),
         State.x = fct_reorder(State.x, -all_farms),
         farm_size = as.factor(farm_size),
         farm_size = fct_collapse(farm_size,
                                  "10-179 acres" = c("10-49 Acres", "50-179 Acres")),
         farm_size = fct_reorder(farm_size, total_farms))%>%
  ggplot(aes(x = State.x, y = total_farms)) +
    geom_col(aes(fill = farm_size)) + 
    labs(x = "State",
         y = "Number of Farms",
         title = "American agriculture is an industry of small farms",
         subtitle = "Most farms are between 10-180 acres nationwide.",
         caption = "Data Source: 2012 Agricultural Census",
         fill = "Farm Size") +
    main_theme +
    theme(axis.text.x = element_text(angle = 90),
          strip.text = element_text(face = "bold", 
                                  size = rel(0.2)),
          legend.position = c(0.55, 0.8))+ 
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values = categorical_cols, aesthetics = "fill")
```

## Food Security and Farms

There is a positive relationship between number of farms in a state and number of 
residents with low food access. Low access to food is defined as living more than one mile
away from a groccery store in urban areas, and more than 10 miles away from a 
groccery store in rural areas. Food insecurity is defined as being unable to 
provide adequate food for one or more household members at some time during 
the year. While food access is related to number of farms, food insecurity is not
related to access to food. There may be food produced in a state, but the ability of 
residents to buy that food is another matter entirely. 

```{r warning=FALSE, message=FALSE, error=FALSE}
byState <- ctyFarms %>%
  group_by(State.x) %>%
  summarise(low_acc_pop = sum(LACCESS_POP10)/1000000,
            farms = sum(DATA1_1),
            snap_redemp = sum(REDEMP_SNAPS12,
                              na.rm = TRUE)/1000000,
            payments = sum(DATA1_33, na.rm = TRUE),
            cropland = sum(DATA1_15, na.rm = TRUE),
            farmland = sum(DATA1_2, na.rm = TRUE),
            avg_land = mean(DATA1_3, na.rm = TRUE),
            obeseAd = mean(PCT_OBESE_ADULTS13),
            insec = mean(VLFOODSEC_10_12 + FOODINSEC_10_12)) %>%
  filter(State.x != "0")
accAss <- right_join(access, assistance, by = "FIPS")
accAssSec <- right_join(accAss, security, by = "FIPS")
ctyFarms <- inner_join(accAssSec, agCensus, by = "FIPS")
ctyFarms <- inner_join(ctyFarms, health, by = "FIPS") %>%
              filter('COUNICP' != 0)

byState %>%
  ggplot(aes(x = farms, y = low_acc_pop)) +
    geom_point(aes(color = insec), 
               alpha = 0.9, 
               size = 7) +
    geom_text(aes(label = State.x),size = 2, color = "white") +
    scale_size_continuous(range = c(3, 8), labels = comma) +
    scale_y_continuous(breaks = pretty(byState$low_acc_pop)) +
    scale_x_log10(breaks = pretty(byState$farms))+
    scale_color_continuous(low = "#e5f5e0", 
                           high = "#006d2c") +
  labs(x = "Number of Farms",
       y = "Residents with Low Food Access (Millions)", 
       color = "Percent Food\nInsecure",
       caption = "Data Source: Agricultural Census 2012",
         title = "Texas is an outlier in number of farms, low food access, and SNAP Redemptions",
         subtitle = "Food security is not related to state food production, but low access is positively related to number of \nfarms.") +
  main_theme +
  theme(legend.position = c(0.2, 0.5), 
        legend.background = element_rect(fill = "transparent"),
        legend.title = element_text(size = 9),
        axis.text.x = element_text(angle = 90)) +
  guides(colour = guide_legend(title.position = "right"),
         size = guide_legend(title.position = "right")) +
  stat_smooth(method="lm", se=FALSE, 
                formula=(y ~ exp(x)),colour="red", linetype = "dashed") 
```

## Farms Over Time

Most states lost farms or stayed constant between 1982 and 2012. Texas is a major
outlier. Not only did the state start with more farms than any other state, but it 
consistently gained farms over the 40 year period.

```{r warning=FALSE, message=FALSE, error=FALSE}
farmbystate <-
farmsTime %>%
  na.omit(farmsTime) %>%
  gather("1950", "1982", "2002", "2007", "2012",
         key = "year",
         value = "farms") %>%
  filter(FIPS != 0 & FIPS != 11) %>%
  mutate(farms = as.integer(farms)) %>%
  group_by(Areaname, year) %>%
  summarise(total_farms = sum(farms, na.rm = TRUE))
farmslate <- farmbystate %>%
  filter(year != 1950)

ggplot(data = farmslate, aes(x= year, y = total_farms)) +
  geom_line(aes(group = Areaname), color = "grey", alpha = 0.6) +
  geom_line(data = filter(farmslate, 
                          Areaname == "Texas"), 
            aes(group = Areaname), 
            color = "#006d2c") + 
  annotate("text", x = 1.5, y = 225000, label = "Texas")+
  labs(x = "Year",
       y = "Number of Farms",
       title = "Texas consistently gained farms from 1982-2012, when \nmost other states lost farms.",
       subtitle = "Texas has had the most farms in the US for 40 years.",
       caption = "Source: Iowa Community Indicators Program") +
  scale_y_continuous(labels = comma) +
  scale_x_discrete(breaks = round(seq(min(farmslate$year), max(farmslate$year), by = 1),1),
                   expand = c(0,.1)) +
  main_theme
```

## Crops

```{r warning=FALSE, message=FALSE, error=FALSE}
crops <- agCensus %>%
  select(COUNICP, NAME,DATA1_16, DATA1_139, DATA1_133, DATA1_64, DATA1_68, DATA1_70, DATA1_94,
         DATA1_103, DATA1_82, DATA1_85, DATA1_88, DATA1_97, DATA1_109, DATA1_112, DATA1_115,
         DATA1_121, DATA1_124, DATA1_127, DATA1_130, DATA1_91) %>%
  filter(COUNICP == 0) %>%
  head(n = 50) 

crop_name <- crops %>%
  gather(DATA1_139, DATA1_133, DATA1_64, DATA1_68, DATA1_70, DATA1_94,
         DATA1_103, DATA1_82, DATA1_85, DATA1_88, DATA1_97, DATA1_109, DATA1_112, DATA1_115,
         DATA1_121, DATA1_124, DATA1_127, DATA1_130, DATA1_91,
         key = "crop", value = "acres") %>%
  mutate(crop = as.factor(crop)) %>%
  mutate(crop = fct_recode(crop,
                            "Orchards"="DATA1_139",
                            "Vegetables"="DATA1_133" ,
                            "Corn" = "DATA1_64",
                            "Silage corn"= "DATA1_68",
                            "Wheat" = "DATA1_70",
                            "Soybeans" = "DATA1_94" ,
                            "Cotton" = "DATA1_103"  ,
                            "Oats" = "DATA1_82" ,
                            "Barley" = "DATA1_85"  ,
                            "Sorghum" = "DATA1_88" ,
                            "Sorghum silage" = "DATA1_91",
                            "Dry Beans" = "DATA1_97"  ,
                            "Tobacco" = "DATA1_109"  ,
                            "Hay" = "DATA1_112"  ,
                            "Rice" = "DATA1_115" ,
                            "Sugar Beets" = "DATA1_121" ,
                            "Sugar Cane" = "DATA1_124"  ,
                            "Pineapples" = "DATA1_127" ,
                            "Peanuts" = "DATA1_130")) %>%
  na.omit() 

crop_name %>%
  mutate(crop = fct_collapse(crop,
                             "Animal Feed" = c("Silage corn",
                                               "Hay",
                                               "Sorghum silage"),
                             "Other" = c("Peanuts",
                                         "Sugar Cane",
                                        "Sugar Beets",
                                        "Dry Beans",
                                        "Tobacco",
                                        "Cotton",
                                        "Oats",
                                         "Barley",
                                         "Sorghum",
                                         "Rice"),
                             "Fruits/Vegetables" = c("Vegetables",
                                                     "Pineapples",
                                                     "Orchards"))) %>%
  select(acres, crop)%>%
  mutate(crop = factor(crop,levels = c("Animal Feed",
                       "Corn", 
                       "Soybeans",
                       "Wheat",
                       "Other",
                       "Fruits/Vegetables"))) %>%
  group_by(crop) %>%
  summarise(tot_acres = sum(acres),
            proportion = tot_acres/389690400) %>%
  ggplot(aes(x = "", y = tot_acres)) +
    geom_bar(stat = "identity", aes(fill = crop)) +
    annotate("text", x = "", y = 5000, label = "2%") +
    annotate("text", x = "", y = 30000000, label = "7%") +
    annotate("text", x = "", y = 60000000, label = "13%") +
    annotate("text", x = "", y = 135000000, label = "20%") +
    annotate("text", x = "", y = 200000000, label = "22%") +
    annotate("text", x = "", y = 350000000, label = "43%") +
    scale_y_continuous(labels = comma)+
    coord_polar("y") +
    scale_fill_manual(values = categorical_cols, aesthetics = "fill") +
    main_theme +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text = element_blank(),
          axis.line = element_blank(),
          legend.position = "right",
          legend.direction = "vertical") +
    labs(title = "The majority of cropland in the US is used to grow animal feed",
         subtitle = "Fruits, vegetables, and all other crops make up less than",
         caption = "Source: 2012 Agricultural Census") 
    
```

